### 2016-06-24 intro.Rmd
ENH:  More responses.

White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences. These lesions are distinguished from prominent perivascular spaces seen on T2-weighted imaging by the lack of fluid suppression of signal on FLAIR sequences. WMHs in the periventricular and deep brain regions are associated with normal aging and neurological conditions including hypertension and stroke. WMHs are also a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric [@Bigler:2013aa;@Smitherman:2016aa] and adult
[@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa]. Further, the regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa]. Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury [@Pierallini:2000aa]. Further, volume of FLAIR lesions within the corpus callosum, brainstem, and thalamus in patients with severe TBI correlates with Glasgow Outcome-Extended (GOS-E) scores [@Moen:2014aa]. Additionally,
in patients who are comatose following severe TBI
the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome [@Weiss:2008aa].

Despite the above findings, outside of multiple sclerosis, WMHs are not routinely employed as a diagnostic measure in clinical practice. Their presence within asymptomatic patients or in association with a variety of conditions, such as stroke, dementia, neuroinflammatory conditions, and TBI challenge their utility in narrowing a radiological differential diagnosis. Further, performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical. Despite the limited inclusion of WMH observations in routine radiological reports, two large meta-analyses demonstrated an association between WMHs, cognitive function, increased risk of stroke, dementia, and death [@Debette:2010aa;@Kloppenborg:2014aa]. As such, the development of automated methods for the rapid identification and quantification of WMHs within individual patients may allow for identification of correlative patterns between WMH number, volume, distribution, and disease state. Further, the development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.

@@

Machine learning and pattern recognition techniques have seen increased application
for various medical image analysis workflows (see, for example, the annual
Workshop on Machine Learning in Medical Imaging held in conjunction with the Medical
Image Computing and Computer-Aided Intervention (MICCAI) international meeting).
Popular techniques such as support vector machines and neural networks have been applied
successfully to clinically relevant imaging tasks such as supervised image segmentation
(e.g., [@Bauer:2011aa]) and diagnostic prediction (e.g., [@Tong:2014aa;@Liu:2013aa]).
Facilitating the current employment of such techniques are the number of available imaging
data sets [@Van-Horn:2014aa] and the public availability of data science packages such as
SciPy [@scipy] and the R project for statistical computing [@R] and their associated
extensions.

The random forests framework [@breiman2001] is a popular machine learning technique that has demonstrated significant utility for supervised segmentation tasks (e.g.,
normal human brain segmentation [@yi2009]) and other computer vision applications
(e.g., human gait detection [@viola2005]).
In the context of neuropathology, random forest-based paradigms have been employed in the
delineation of multiple sclerosis lesions [@geremia2011], stroke lesions [@Pustina:2016aa],
and brain tumors [@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa] for both
single and multi-modal acquisition protocols.  Of note, these
latter random forest approaches for brain tumor segmentation have performed well in recent
international competitions established in response to the lack of objective comparisons between segmentation algorithms (i.e.,  the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa].

@@

In this work, we develop a concatenated random forest framework with a tailored contextual
feature image set (both spatial and intensity-based)
for segmenting WMHs in a large TBI cohort.  Additionally,
the entire framework is provided publicly through the well-known open-source
Advanced Normalization Tools (ANTs)[^i1] and ANTsR[^i2] toolkits.
Further motivating this research is the availability of several large publicly available  imaging data sets that permits testing reproducibility of this automated routine for WMH segmentation and quantification.

>>>>

\textcolor{blue}{White matter hyperintensities (WMHs) are foci of abnormally increased signal intensity seen within white matter regions within the cerebrum and brainstem on fluid attenuation inversion recovery (FLAIR) magnetic resonance imaging (MRI) sequences.
WMHs are a frequent finding following traumatic brain injury (TBI) and have been correlated with functional outcome and injury severity in both pediatric} [@Bigler:2013aa;@Smitherman:2016aa] \textcolor{blue}{and adult} [@Marquez-de-la-Plata:2007aa;@Moen:2014aa;@Ding:2008aa;@Pierallini:2000aa] \textcolor{blue}{populations.}

\textcolor{blue}{Further research involving WMHs has shown that regional distribution and volume of WMHs have been shown to possess prognostic value in the TBI patient} [@Pierallini:2000aa;@Weiss:2008aa;@Smitherman:2016aa;@Levin:1988aa]\textcolor{blue}{. Specifically, lesion volume in corpus callosum correlates with functional scores in the acute phase following injury, while lesion volume in frontal lobes correlates with scores at 1 year following injury} [@Pierallini:2000aa]\textcolor{blue}{.
Further, volume of FLAIR lesions within the corpus
callosum, brainstem, and thalamus in patients with severe TBI correlates
with Glasgow Outcome-Extended (GOS-E) scores---a numeric groupwise assessment used to
classify ``outcome'' in TBI patients where ``outcome'' refers to the
spectrum of possible prognoses from death to disability to recovery.} [@Moen:2014aa]\textcolor{blue}{. Additionally,
in patients who are comatose following severe TBI
the regional distribution of FLAIR lesions within the pons, midbrain, hypothalamus, basal forebrain, parietal, temporal, occipital lobes, and insula along with the observation of grasping or chewing behavior are associated with poor outcome} [@Weiss:2008aa].

\textcolor{blue}{Despite the above findings which demonstrate that WMHs have potential prognostic value, they are not routinely employed as a diagnostic measure in clinical practice.  Performing a comprehensive manual counting of number and distribution of lesions in the clinical setting is simply not practical. As such, the development of automated methods for the rapid identification and quantification of WMHs within individual patients may allow for identification of correlative patterns between WMH number, volume, distribution, and disease state. Further, the development of such lesion quantification approaches may allow for the practical inclusion of this type of information within routine radiological practice.  In this work, we present an automated framework for quantification of WMHs in
multi-modal MRI using the random forest machine learning technique.}

@@

<!--
Machine learning and pattern recognition techniques have seen increased application
for various medical image analysis workflows (see, for example, the annual
Workshop on Machine Learning in Medical Imaging held in conjunction with the Medical
Image Computing and Computer-Aided Intervention (MICCAI) international meeting).
Popular techniques such as support vector machines and neural networks have been applied
successfully to clinically relevant imaging tasks such as supervised image segmentation
(e.g., [@Bauer:2011aa]) and diagnostic prediction (e.g., [@Tong:2014aa;@Liu:2013aa]).
Facilitating the current employment of such techniques are the number of available imaging
data sets [@Van-Horn:2014aa] and the public availability of data science packages such as
SciPy [@scipy] and the R project for statistical computing [@R] and their associated
extensions.
-->

The random forests framework [@breiman2001] is a popular machine learning technique that has demonstrated significant utility for supervised segmentation tasks (e.g.,
normal human brain segmentation [@yi2009]) and other computer vision applications
(e.g., [@viola2005]).
\textcolor{blue}{Random forest-based paradigms have been successfully employed in the
delineation of other neuropathologies} [@geremia2011;@Pustina:2016aa;@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa]
\textcolor{blue}{for both single and multi-modal acquisition protocols.}

<!--
Of note, these
latter random forest approaches for brain tumor segmentation have performed well in recent
international competitions established in response to the lack of objective comparisons between segmentation algorithms (i.e.,  the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa].
-->

@@

In this work, we develop a concatenated random forest framework with a
feature image set (both spatial and intensity-based)
for segmenting WMHs in a large TBI cohort.
The entire framework is built on the well-known open-source
Advanced Normalization Tools (ANTs)[^i1] and ANTsR[^i2] toolkits.
Further motivating this research is the availability of several large publicly available  imaging data sets that permits testing reproducibility of this automated routine for WMH segmentation and quantification.

### 2016-06-24 methods.Rmd
ENH:  More responses.

@@

Crucial to these supervised segmentation approaches are the creation and selection of
"features" as input in conjunction with
expertly identified structures of interest for model construction.  For the targeted application in this work (i.e., WMHs),
tissue classification is performed at the voxelwise level.  In other words, each voxel
within the region of interest is sent through the ensemble of decision trees and receives
a set of classification votes from each tree thus permitting a regression or classification
solution.  Since this procedure is performed at the voxelwise level, intensity information
alone is insufficient for good segmentation performance due to the lack of spatial context.
For example, as pointed out in [@Neema:2009aa], higher intensities can be found at the
periventricular caps in normal subjects which often confounds automated lesion detection
algorithms.  Other potential confounds include MR signal inhomogeneity and noise.  Therefore,
even though machine learning and pattern recognition techniques are extremely powerful and
have significant potential, just as crucial to outcome is the creative construction and
deployment of salient feature images which we detail below.

@@

![Representation of Stage 1 feature images for subject 01C1019.  The
FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
[@Avants:2014aa] to the space of the T1 image.  The three modality images are then preprocessed
(N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
the MMRR symmetric template and corresponding priors [@Tustison:2014ab] applied to the T1 image.
The feature images are then generated for voxelwise input to the RF model which results in
the voting maps illustrated on the right.  This gives a probabilistic classification of
tissue type.  Not shown are the probability and voting images for the brain stem and
cerebellum.](Figures/featureImages.png)

@@

As mentioned previously, input for each subject comprises FLAIR, T1-, and T2-weighted
acquisitions.  The FLAIR and T2 images are rigidly registered to the T1 image using
the open-source Advanced Normalization Tools (ANTs) [@Avants:2014aa].  The aligned images
are then preprocessed using the denoising algorithm of [@Manjon:2010aa] followed by
N4 bias correction [@Tustison:2010ac] which are then normalized to the intensity
range $[0,1]$.  Although we could have used an alternative intensity standardization
algorithm (e.g., [@nyul2000]), we found that a simple linear rescaling produced better results similar to previous work[@Tustison:2015aa].

@@

The T1 image is then processed via the ANTs brain
extraction and normal tissue segmentation pipelines [@Tustison:2014ab].
Any effect of pathology is minimal because WMHs are not commonly seen in T1-weighted MRI.  This protocol involves preprocessing using N4 bias correction followed by a template-based strategy for brain extraction.  Once the
brain has been extracted, we apply a Bayesian-based segmentation algorithm
with a template-based prior probability strategy to segment the
parenchymal tissue types.  The result is a mask delineating the brain parenchyma and probabilistic estimates of the
CSF, gray matter, white matter, deep gray matter, brain stem, and
cerebellum.  These provide the expertly annotated labels for the first six tissue labels
given above.  The WMHs were manually identified by one of the authors
(J. R. S.) using the ITK-SNAP tool [@Yushkevich:2006aa].  Segmentation is performed using
the ANTs Atropos tool [@Avants:2011aa] and
multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public
MMRR data set  [@landman2011] (cf Figure 3).

![Canonical views of the mutlivariate, bilaterally symmetric template constructed
from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
the components relevant for this work).  Template construction is detailed in
[@Tustison:2015aa].  These images are important for specific intensity-based
features.](Figures/MMRR.png)

To model the intensity information the first set of images simply includes the
preprocessed and normalized intensity FLAIR, T1, and T2 image voxel values.  We also calculate a set
of neighborhood statistics (mean, standard deviation, and skewness) feature images using
a Manhattan radius of one voxel given the typical size
of individual WMHs. For each of the normalized
images, we calculate the difference in intensities with the corresponding warped template
component.  Previous success in the international brain tumor segmentation
competition [@Menze:2015aa] was based on an important set of intensity features that were created
from multi-modal templates mentioned previously [@Tustison:2015aa] and listed in Table 1.  We employ the
same strategy here.  For example, the template
difference feature image for the FLAIR image, $S_{FLAIR}$ is calculated as:

@@

is the transform
which maps from the individual subject space to the
template space and $T_{FLAIR}$ is the FLAIR template component.  Also,
to take advantage of the gross bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that WMHs do not generally manifest symmetrically across hemispheres,
we use the symmetric templates to compute the contralateral intensity
differences as an additional intensity feature.  For the FLAIR component, this contralateral
difference image is calculated from

@@

where $\phi_R$ denotes a horizontal reflection perpendicular to the mid-sagittal plane of
the symmetric template.

The segmentation probability images described above are used as feature images to provide a
spatial context for the random forest model prediction step.  Additional spatial contextual
feature images include the distance maps [@maurer2003] based on the csf, gray matter, and
deep gray matter images.  These latter images are intended to help distinguish white matter
hyperintensities from false positives induced by the partial voluming at the gray/white
matter interface.  A third set of images are based on the voxel location within the
space of the template.  The T1 image of the subject is registered to the T1 template
component using a B-spline variant [@Tustison:2013ac] of the well-known ANTs Symmetric
Normalization (SyN) algorithm [@Avants:2011ab].  Since the inverse transform is also
derived as part of the registration process, we can warp the voxel index locations
back to the space of the individual subject which motivates similar work by others [@Anbeek:2004aa].  However, this previous work lacks the normalization to the standard
coordinate system provided by the template to dramatically improve spatial specificity
across all subjects.

@@

In previous brain tumor segmentation work [@Tustison:2015aa], it was demonstrated that a
concatenated supervised approach, whereby the prediction output from the first random forest
model serves as partial input for a second random forest model, can significantly improve
segmentation performance.  We do the same thing for the work described here where we employ two stacked random forests (or two "stages").  The Stage 1
feature images of the training data (as described previously) are used to construct the
Stage 1 model.  The training data Stage 1 features are then used to produce the voxelwise
"voting maps" (i.e., the classification count of each decision tree for each tissue label) via the Stage 1 random forest model.  All the Stage 1 features plus the Stage 1 voting maps
are used as input to the Stage 2 model.  In addition, we use the Stage 1 voting maps as
tissue priors (i.e., probabilistic estimates of the tissue spatial locations) for a second application of the Atropos maximum aposteriori algorithm with an
additional Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].  However, for
the second stage we use all three aligned preprocessed images for a multivariate
segmentation.  The resulting seven posterior probability images constitute a third additional feature image set for Stage 2.

### Code and data availability

As pointed out in a recent comprehensive multiple sclerosis lesion segmentation review
[@Garcia-Lorenzo:2013aa], although the number of algorithms reported in the literature
is quite extensive, there were only four publicly available segmentation algorithms
at the time of writing this article.  In contrast to the current work, none are based on supervised learning.  As we did for
our brain tumor segmentation algorithm [@Tustison:2015aa], all of the code described in
this work is publicly available through the open-source ANTs/ANTsR toolkits.  Through
ANTsR (an add-on toolkit which, in part, bridges ANTs and the R statistical project) we
use the _randomForest_ package [@liaw2002] using the default settings with
2000 trees per model and 500 randomly selected samples per label per image.
Note that we saw little variation in performance when these parameters were changed (i.e. up to 1000 random samples and as little as 1000 trees) which is consistent with our previous experience.

In addition, similar to our previous offering [^1], we plan on creating
a self-encapsulated example to showcase the proposed methodology.  The fact that
the data will also be made available through the Federal Interagency Traumatic Brain Injury Research (FITBIR) repository along
with the manual labelings will facilitate reproducibility on the part of the reader
as well as any interest in extending the proposed framework to other data sets.

[^1]: https://github.com/ntustison/ANTsAndArboles

@@

In order to evaluate the protocol described, we performed a leave-one-out evaluation using
the data acquired from the 24 subjects described above.  Initial processing included the creation
of all Stage 1 feature images for all subjects.  The initial brain segmentation of each T1 image and the
manual white matter hyperintensity tracings were combined to provide the truth labels
for the training data.  The "truth" labels are the seven anatomical regions given above.

>>>>

\textcolor{blue}{The first author (J. R. S.) performed the manual WMH tracings for all 26
subjects.  J. R. S. is a radiologist certified by the American Board of Radiology, with a certificate of advanced qualification in vascular and interventional radiology, over 18 years of research experience in TBI, and 6 years of clinical imaging experience.  All
multi-modal MR dicom image slices were converted to the nifti file format} [^1].
\textcolor{blue}{Given the
utility of the FLAIR sequence in detecting white matter lesions in TBI}  [@Marquez-de-la-Plata:2007aa]\textcolor{blue}{, all nifti image volumes for each subject were rigidly aligned to the FLAIR image of that subject using the ANTs software} [@Avants:2014aa] \textcolor{blue}{.  The normalized MRI volumes were then provided to
J. R. S. who traced each lesion using the ITK-SNAP tool} [@Yushkevich:2006aa]
\textcolor{blue}{
which has multi-image overlay capabilities for visualizing all modalities in all three
canonical views.}

[^1]: http://nifti.nimh.nih.gov/nifti-1

@@

\textcolor{blue}{
Crucial to these supervised segmentation approaches are the creation and selection of
``features'' as input (i.e., feature images constructed from the training data)
in conjunction with expertly identified structures of interest
(i.e., WMHs) for model construction.}  For the targeted application in this work,
tissue classification is performed at the voxelwise level.  In other words, each voxel
within the region of interest is sent through the ensemble of decision trees and receives
a set of classification votes from each tree thus permitting a regression or classification
solution.  Since this procedure is performed at the voxelwise level, intensity information
alone is insufficient for good segmentation performance due to the lack of spatial context.
For example, as pointed out in [@Neema:2009aa], higher intensities can be found at the
periventricular caps in normal subjects which often confounds automated lesion detection
algorithms.  Other potential confounds include MR signal inhomogeneity and noise.  Therefore,
even though machine learning and pattern recognition techniques are extremely powerful and
have significant potential, just as crucial to outcome is the creative construction and
deployment of salient feature images which we detail below.

@@

![Representation of Stage 1 feature images for subject 01C1019.  The
FLAIR, T1-, and T2-weighted images are rigidly pre-aligned
[@Avants:2014aa] to the space of the FLAIR image.  The three modality images are then preprocessed
(N4 bias correction [@Tustison:2010ac] and adaptive denoising [@Manjon:2010aa]) followed by
application of standard ANTs brain extraction and $n$-tissue segmentation protocols using
the MMRR symmetric template and corresponding priors [@Tustison:2014ab] applied to the T1 image.
The feature images are then generated for voxelwise input to the RF model which results in
the voting maps illustrated on the right.  This gives a probabilistic classification of
tissue type.  Not shown are the probability and voting images for the brain stem and
cerebellum.](Figures/featureImages.png)

@@

As mentioned previously, input for each subject comprises FLAIR, T1-, and T2-weighted
acquisitions.  The T1 and T2 images are rigidly registered to the FLAIR image using
the open-source Advanced Normalization Tools (ANTs) [@Avants:2014aa].  The aligned images
are then preprocessed using the denoising algorithm of [@Manjon:2010aa] followed by
N4 bias correction [@Tustison:2010ac] which are then normalized to the intensity
range $[0,1]$.  Although we could have used an alternative intensity standardization
algorithm (e.g., [@nyul2000]), we found that a simple linear rescaling produced better results similar to previous work[@Tustison:2015aa].

@@

The T1 image is then processed via the ANTs brain
extraction and normal tissue segmentation pipelines [@Tustison:2014ab].
  The result is a mask delineating the brain parenchyma and probabilistic estimates of the
CSF, gray matter, white matter, deep gray matter, brain stem, and
cerebellum [@Avants:2011aa].  These provide the expertly annotated labels for the
first six tissue labels given above.  Tissue prior probability maps for segmentation
are from multi-model optimal symmetric shape/intensity templates [@Tustison:2015aa] created from the public MMRR data set  [@landman2011] (cf Figure 3).

<!--
![Canonical views of the mutlivariate, bilaterally symmetric template constructed
from the MMRR data set [@landman2011] (only shown are the FLAIR, T1, and T2 modalities---
the components relevant for this work).  Template construction is detailed in
[@Tustison:2015aa].  These images are important for specific intensity-based
features.](Figures/MMRR.png)
-->

Feature values include the preprocessed FLAIR, T1, and T2 image voxel intensities.
We also calculate a set of neighborhood statistics (mean, standard deviation, and skewness) feature images using
a Manhattan radius of one voxel given the typical size
of individual WMHs. For each of the preprocessed
images, we calculate the difference in intensities with the corresponding warped template
component.  Previous success in the international brain tumor segmentation
competition [@Menze:2015aa] was based on an important set of intensity features that were created
from multi-modal templates mentioned previously [@Tustison:2015aa] and listed in Table 1.  We employ the
same strategy here.

<!--
For example, the template
difference feature image for the FLAIR image, $S_{FLAIR}$ is calculated as:

@@

is the transform
which maps from the individual subject space to the
template space and $T_{FLAIR}$ is the FLAIR template component.
-->

To take advantage of the gross bilateral symmetry of the normal brain (in terms of both shape and intensity), and the fact that WMHs do not generally manifest symmetrically across hemispheres,
we use the symmetric templates to compute the contralateral intensity
differences as an additional intensity feature.

<!--
For the FLAIR component, this contralateral
difference image is calculated from

@@

where $\phi_R$ denotes a horizontal reflection perpendicular to the mid-sagittal plane of
the symmetric template.
-->

The segmentation probability images described above are used as feature images to
provide a spatial context for the random forest model prediction step.  Additional spatial contextual
feature images include the distance maps [@maurer2003] based on the csf, gray matter, and
deep gray matter images.  These latter images are intended to help distinguish white matter
hyperintensities from false positives induced by the partial voluming at the gray/white
matter interface.  A third set of images are based on the voxel location within the
space of the template.
\textcolor{blue}{Similar feature images were used in}
[@Anbeek:2004aa]
\textcolor{blue}{although, unlike the proposed framework, this previous work lacks
normalization to the standard
coordinate system provided by the template to dramatically improve spatial specificity
across all subjects.  To generate these images, the T1 image of each subject is
registered to the T1 template component using a B-spline variant} [@Tustison:2013ac]
\textcolor{blue}{of the well-known ANTs Symmetric Normalization (SyN) algorithm} [@Avants:2011ab].
\textcolor{blue}{Using the derived transforms, the template coordinate images are warped back to the space of the individual subject.}

@@

In previous brain tumor segmentation work [@Tustison:2015aa], it was demonstrated that a
concatenated supervised approach, whereby the prediction output from the first random forest
model serves as partial input for a second random forest model, can significantly improve
segmentation performance.  We do the same thing for the work described here where we employ two stacked random forests (or two "stages").  The Stage 1
feature images of the training data (as described previously) are used to construct the
Stage 1 model.  The training data Stage 1 features are then used to produce the voxelwise
"voting maps" (i.e., the classification count of each decision tree for each tissue label) via the Stage 1 random forest model.  All the Stage 1 features plus the Stage 1 voting maps
are used as input to the Stage 2 model.  In addition, we use the Stage 1 voting maps as
tissue priors (i.e., probabilistic estimates of the tissue spatial locations) for a second application of the $6$-tissue segmentation algorithm with an
additional Markov Random Field spatial prior (MAP-MRF) [@Avants:2011aa].
\textcolor{blue}{In order to maximize the spatial information for the $n$-tissue segmentation process following the voxelwise RF classification of Stage 1, we use
all three aligned preprocessed images for multivariate segmentation during the
second stage.}
The resulting seven posterior probability images constitute a third additional feature image set for Stage 2.

### Implementation

As pointed out in a recent comprehensive lesion segmentation review
[@Garcia-Lorenzo:2013aa], although the number of algorithms reported in the literature
is quite extensive, there were only four publicly available segmentation algorithms
at the time of writing this article.  In contrast to the current work, none are based on supervised learning.  As we did for
our brain tumor segmentation algorithm [@Tustison:2015aa], all of the code described in
this work is publicly available through the open-source ANTs/ANTsR toolkits.  Through
ANTsR (an add-on toolkit which, in part, bridges ANTs and the R statistical project) we
use the _randomForest_ package [@liaw2002] using the default settings with
2000 trees per model and 500 randomly selected samples per label per image.
Note that we saw little variation in performance when these parameters were changed (i.e. up to 1000 random samples and as little as 1000 trees) which is consistent with our previous experience.

In addition, similar to our previous offering [^2], we plan on creating
a self-encapsulated example to showcase the proposed methodology
\textcolor{blue}{which will also be available on github.}[^3]  The fact that
the data will also be made available through the Federal Interagency Traumatic Brain Injury Research (FITBIR) repository along
with the manual labelings will facilitate reproducibility on the part of the reader
as well as any interest in extending the proposed framework to other data sets.

[^2]: https://github.com/ntustison/ANTsAndArboles

[^3]: https://github.com/ntustison/WatchMeHyperventilate

@@

In order to evaluate the protocol described, we performed a leave-one-out evaluation using
the data acquired from the 24 subjects described above.  Initial processing included the creation of all Stage 1 feature images for all subjects.  The initial brain segmentation of each T1 image and the manual white matter hyperintensity tracings were combined to provide the truth labels for the training data.  The ``truth'' labels are the seven anatomical regions given above.

### 2016-06-24 results.Rmd
ENH:  More responses.

One can also use these measurements as a type of sanity check.  For example, from the Stage
1  plot, one can see that the _MeanDecreaseAccuracy_ values for the location indices in the
anterior-posterior direction (i.e., _TemplateIndicesWarped1_) are greater than
those for either the inferior-superior (i.e., _TemplateIndicesWarped2_) or the
left-right (i.e., _TemplateIndicesWarped0_) directions in the space of the symmetric
template.

@@

In Figure 7 are the segmentation comparisons derived from manual segmentations of the same
data.  Despite the large variability characteristic with manual labelings in related fields
[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
current clinical practices and the methodology proposed herein is readily adapted to refinements
in training data.  On the left of Figure 7 are the improvement in Dice values [@tustison2009],
i.e.,

$$ Dice = 2 \frac{\sum_r|S_r \cap T_r|}{\sum_r|S_r| + |T_r] } $$

over all white matter hyperintensities when comparing the segmentations between the two stages
where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
represents the intersection between the manual/automated lesion pair.  Performing
the second round of supervised learning improves these Dice values.  One can also note from the
right side of Figure 6 that the total lesion load volume illustrates a few subjects that are
severe outliers in terms of the number of false positives.  The second round helps to
correct this issue.

>>>>

<!--
One can also use these measurements as a type of sanity check.  For example, from the Stage
1  plot, one can see that the _MeanDecreaseAccuracy_ values for the location indices in the
anterior-posterior direction (i.e., _TemplateIndicesWarped1_) are greater than
those for either the inferior-superior (i.e., _TemplateIndicesWarped2_) or the
left-right (i.e., _TemplateIndicesWarped0_) directions in the space of the symmetric
template.
-->

@@

In Figure 7 are the segmentation comparisons derived from manual segmentations of the same
data.  Despite the large variability characteristic with manual labelings in related fields
[@Grimaud:1996aa;@styner2008;@Garcia-Lorenzo:2013aa], such labelings are characteristic of
current clinical practices and the methodology proposed herein is readily adapted to refinements
in training data.

<!--
i.e.,

$$ Dice = 2 \frac{\sum_r|S_r \cap T_r|}{\sum_r|S_r| + |T_r] } $$
-->

On the left of Figure 7 are the improvement in Dice values [@tustison2009],
over all white matter hyperintensities when comparing the segmentations between the two stages
where the sum is taken over all individually labeled manual, $T_r$, and automated, $S_r$, lesions and $\cap$
represents the intersection between the manual/automated lesion pair.  Performing
the second round of supervised learning improves these Dice values.  One can also note from the
right side of Figure 6 that the total lesion load volume illustrates a few subjects that are
severe outliers in terms of the number of false positives.  The second round helps to
correct this issue.
